---
output: html_document
editor_options: 
  chunk_output_type: console
---
White Wine Quality by Abbas A. Alali
========================================================

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo=FALSE)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using in your analysis in this code
# chunk.
#install.packages("corrplot", dependencies=TRUE, repos='http://cran.rstudio.com/')
suppressMessages(library(ggplot2))
suppressMessages(library(gridExtra))
suppressMessages(library(corrplot))
suppressMessages(library(memisc))
suppressMessages(library(RColorBrewer))
```

> The dataset that i will be using in this exploratory data analysis process\
>is the White Wine dataset which is public available for research. 
>
>The details are described in [Cortez et al., 2009].
>
>     P. Cortez, A. Cerdeira, F. Almeida, T. Matos and J. Reis. Modeling wine\
>     preferences by data mining from physicochemical properties. \
>     In Decision Support Systems, Elsevier,47(4):547-553. ISSN: 0167-9236.
>
>     Available at: 
>
>     [@Elsevier](http://dx.doi.org/10.1016/j.dss.2009.05.016)
>                
>     [Pre-press (pdf)](http://www3.dsi.uminho.pt/pcortez/winequality09.pdf)
>                
>     [bib](http://www3.dsi.uminho.pt/pcortez/dss09.bib)
>
> The dataset used has been downloaded from [this link](https://www.google.com/url?q=https://s3.amazonaws.com/udacity-hosted-down\
loads/ud651/wineQualityWhites.csv&sa=D&ust=1518211607164000&usg=AFQjCNGqcmTIF2\
an-RHraAjbtMBpAhsdTg)



```{r message=F, warning=F}
# Load the Data

setwd("C:/Users/Abbas/Documents/DAND/Term 2/Project 2")
wdf = read.csv('wineQualityWhites.csv')

```



# Univariate Plots Section


##***Preliminary Exploration:***

```{r message=F, warning=F}

dim(wdf)

str(wdf)

summary(wdf)

```


###***Quality:***

**The quality of wine varies from 3 to 9 in our dataset. There are no 0 or 10 \
rated wine in the dataset, the minimum rated wine is 3 and the maximum is 9.\
The median quality is 6**

```{r message=F, warning=F}

table(wdf$quality) # get the rating levels

summary(wdf$quality) # summary of rating

wdf$quality.rating = factor(wdf$quality) # store factors into quality.rating variable


ggplot(aes(x = quality.rating), data = wdf) + 
  geom_histogram(stat = "count", fill =I('#2c7fb8')) + # bar polt of quality.rating
  ggtitle("Count of Wine Quality by Rating") + 
  theme(plot.title = element_text(hjust = 0.5)) + # set title to be at center
  xlab('Quality Rating') +
  ylab('Count')

```


**From the summary of the ratings and the plot above, we can see that there are\
very few wine records that are rated as 3 or 9 (25 records out of the \
whole dataset).These records might be excluded from future analysis and plots \
as they will be outliers.**

**Also, a new variable has been created to bucket the rating as Low(3-4), \
Medium(5-7), high(8-9)**

```{r ,message=F, warning=F}

wdf$rating <- cut(wdf$quality, breaks=c(2,4,7,9), labels=c("Low","Medium","High")) #create a bucketed bins of quality
  # 3-4 --> Low
  # 5-7 --> Medium
  # 8-9 --> High

```

###***Fixed Acidity:***

```{r message=F, warning=F}

summary(wdf$fixed.acidity) # summary of fixed acidity

bplot = ggplot(aes(x = fixed.acidity), data = wdf) + 
  geom_bar(stat = "count", fill =I('#2c7fb8')) + # bar polt of fixed.acidity
  ggtitle("Count of Wine by Fixed Acidity Levels (Basic Plot)") + 
  theme(plot.title = element_text(hjust = 0.5)) + # set title to be at center
  xlab('Fixed Acidity') +
  ylab('Count')

rplot = ggplot(aes(x = fixed.acidity), data = wdf) + 
  geom_bar(stat = "count", fill =I('#2c7fb8')) + # bar polt of fixed.acidity
  scale_x_continuous( limit = c(4,9), breaks = seq(4, 9, by = 0.5) ) +
  ggtitle("Count of Wine by Fixed Acidity Levels (Revised Plot)") + 
  theme(plot.title = element_text(hjust = 0.5)) + # set title to be at center
  xlab('Fixed Acidity') +
  ylab('Count')

grid.arrange(bplot, rplot)

```

**The fixed acidity plot looks normally distributed with mean value of (6.855).\
After several adjustments to the x-axis scale, most of the wines are ranged from\
5.5 to 8.5 fixed acidity level**

###***Volatile Acidity:***

```{r message=F, warning=F}

summary(wdf$volatile.acidity) # summary of volatile acidity

bplot = ggplot(aes(x = volatile.acidity), data = wdf) + 
  geom_bar(stat = "count", fill =I('#2c7fb8')) + # bar polt of volatile.acidity
  ggtitle("Count of Wine by Volatile Acidity Levels (Basic Plot)") + 
  theme(plot.title = element_text(hjust = 0.5)) + # set title to be at center
  xlab('Volatile Acidity') +
  ylab('Count')

rplot = ggplot(aes(x = volatile.acidity), data = wdf) + 
  geom_bar(stat = "count", fill =I('#2c7fb8')) + # bar polt of volatile.acidity
  scale_x_continuous( limit= c(0.1, 0.5), breaks = seq(0.1, 0.5, by = 0.05) ) + 
  ggtitle("Count of Wine by Volatile Acidity Levels (Revised Plot)") + 
  theme(plot.title = element_text(hjust = 0.5)) + # set title to be at center
  xlab('Volatile Acidity') +
  ylab('Count')

grid.arrange(bplot, rplot)

```

**With the basic plot of the volatile acidity levels, I can see that most of \
the wine records is ranged from 0.08 to 1.1 acidity levels with (0.2782) \
mean value . After several adjustments to the x-axis scale, most of the wines\
are ranged from 0.15 to 0.40 volatile acidity level**


###***Citric Acid:***

```{r message=F, warning=F}

summary(wdf$citric.acid) # summary of citric acid attribute

bplot = ggplot(aes(x = citric.acid), data = wdf) + 
  geom_bar(stat = "count", fill =I('#2c7fb8')) + # bar polt of citric acid
  ggtitle("Count of Wine by Citric Acid Levels (Basic Plot)") + 
  theme(plot.title = element_text(hjust = 0.5)) + # set title to be at center
  xlab('Citric Acid') +
  ylab('Count')

rplot = ggplot(aes(x = citric.acid), data = wdf) + 
  geom_bar(stat = "count", fill =I('#2c7fb8')) + # bar polt of Citric Acid
  scale_x_continuous( limit = c(0, 0.6), breaks = seq(0, 0.6, by = 0.05) ) + 
  ggtitle("Count of Wine by Citric Acid Levels (Revised Plot)") + 
  theme(plot.title = element_text(hjust = 0.5)) + # set title to be at center
  xlab('Citric Acid') +
  ylab('Count')

grid.arrange(bplot, rplot)

```

**The basic plot of the citric acidity shows that the wine records is ranged\
from 0 to 1.6 acidity levels with mean value of (0.3342). After several\
adjustments to the x-axis scale, most of the wines are ranged from 0.15 to 0.50\
citric acidity level**

###***Residual Sugar:***

```{r message=F, warning=F}

summary(wdf$residual.sugar) # summary of residual sugar attribute

bplot = ggplot(aes(x = residual.sugar), data = wdf) + 
  geom_bar(stat = "count", fill =I('#2c7fb8'), width = 0.1) + # bar polt of residual.sugar
  ggtitle("Count of Wine by Residual Sugar Levels (Basic Plot)") + 
  theme(plot.title = element_text(hjust = 0.5)) + # set title to be at center
  xlab('Residual Sugar') +
  ylab('Count')

rplot = ggplot(aes(x = residual.sugar), data = wdf) + 
  geom_bar(stat = "count", fill =I('#2c7fb8'), width = 0.1) + # bar polt of residual.sugar
  scale_x_continuous( limit = c(0, 20), breaks = seq(0, 20, by = 1) ) + 
  ggtitle("Count of Wine by Residual Sugar Levels (Revised Plot)") + 
  theme(plot.title = element_text(hjust = 0.5)) + # set title to be at center
  xlab('Residual Sugar') +
  ylab('Count')

grid.arrange(bplot, rplot)

```

**The basic plot of the residual sugar shows that the wine records is ranged \
from 0.6 to 65.8 residual sugar levels (which is considered to be a very wide range)\
with mean value of (6.391). After several adjustments to the x-axis scale, \
most of the wines are ranged from 0.5 to 20 residual sugar level**


###***Chlorides:***

```{r message=F, warning=F}

summary(wdf$chlorides) # summary of residual sugar attribute

bplot = ggplot(aes(x = chlorides), data = wdf) + 
  geom_bar(stat = "count", fill =I('#2c7fb8')) + # bar polt of chlorides
  ggtitle("Count of Wine by Chlorides Levels (Basic Plot)") + 
  theme(plot.title = element_text(hjust = 0.5)) + # set title to be at center
  xlab('Chlorides') +
  ylab('Count')

rplot = ggplot(aes(x = chlorides), data = wdf) + 
  geom_bar(stat = "count", fill =I('#2c7fb8')) + # bar polt of chlorides
  scale_x_continuous(limit = c(0, 0.08), breaks = seq(0, 0.08, by = 0.01)) + 
  ggtitle("Count of Wine by Chlorides Levels (Revised Plot)") + 
  theme(plot.title = element_text(hjust = 0.5)) + # set title to be at center
  xlab('Chlorides') +
  ylab('Count')

grid.arrange(bplot, rplot)

```

**The basic plot of the chlorides shows that the wine records is ranged \
from 0 to 0.346 chlorides levels with mean value of (0.04577). After several \
adjustments to the x-axis scale, most of the wines are ranged \
from 0.025 to 0.065 chlorides level**

###***Free Sulfur Dioxide:***

```{r message=F, warning=F}

summary(wdf$free.sulfur.dioxide) # summary of free.sulfur.dioxide attribute

bplot = ggplot(aes(x = free.sulfur.dioxide), data = wdf) + 
  geom_bar(stat = "count", fill =I('#2c7fb8'), width = 1) + # bar polt of free.sulfur.dioxide
  ggtitle("Count of Wine by Free Sulfur Dioxide Levels (Basic Plot)") + 
  theme(plot.title = element_text(hjust = 0.5)) + # set title to be at center
  xlab('Free Sulfur Dioxide') +
  ylab('Count')

rplot = ggplot(aes(x = free.sulfur.dioxide), data = wdf) + 
  geom_bar(stat = "count", fill =I('#2c7fb8')) + # bar polt of free.sulfur.dioxide
  scale_x_continuous(limit = c(0, 70), breaks = seq(0, 70, by = 5)) + 
  ggtitle("Count of Wine by Free Sulfur Dioxide Levels (Revised Plot)") + 
  theme(plot.title = element_text(hjust = 0.5)) + # set title to be at center
  xlab('Free Sulfur Dioxide') +
  ylab('Count')

grid.arrange(bplot, rplot)

```

**The basic plot of the free sulfur dioxide shows that the wine records is\
ranged from 2 to 289 free sulfur dioxide levels with mean value of (35.31). \
After several adjustments to the x-axis scale, most of the wines are ranged\
from 10 to 60 free sulfur dioxide level**

###***Total Sulfur Dioxide:***

```{r message=F, warning=F}

summary(wdf$total.sulfur.dioxide) # summary of free.sulfur.dioxide attribute

bplot = ggplot(aes(x = total.sulfur.dioxide), data = wdf) + 
  geom_bar(stat = "count", fill =I('#2c7fb8'), width = 1) + # bar polt of total.sulfur.dioxide
  ggtitle("Count of Wine by Total Sulfur Dioxide Levels (Basic Plot)") + 
  theme(plot.title = element_text(hjust = 0.5)) + # set title to be at center
  xlab('Total Sulfur Dioxide') +
  ylab('Count')

rplot = ggplot(aes(x = total.sulfur.dioxide), data = wdf) + 
  geom_bar(stat = "count", fill =I('#2c7fb8'),width = 1) + # bar polt of total.sulfur.dioxide
  scale_x_continuous(limit = c(50, 240), breaks = seq(50, 250, by = 10)) + 
  ggtitle("Count of Wine by Total Sulfur Dioxide Levels (Revised Plot)") + 
  theme(plot.title = element_text(hjust = 0.5)) + # set title to be at center
  xlab('Total Sulfur Dioxide') +
  ylab('Count')

grid.arrange(bplot, rplot)

```

**The basic plot of the total sulfur dioxide shows that the wine records is \
ranged from 9 to 440 total sulfur dioxide levels (which is considered to be\
a very wide range) with mean value of (138.4). After several adjustments to \
the x-axis scale, most of the wines are ranged from 60 to 220 total sulfur \
dioxide level**

###***Density:***

```{r message=F, warning=F}

summary(wdf$density) # summary of density

bplot = ggplot(aes(x = density), data = wdf) + 
  geom_bar(stat = "count", fill =I('#2c7fb8'), width = 0.001) + # bar polt of density
  ggtitle("Count of Wine by Density Levels (Basic Plot)") + 
  theme(plot.title = element_text(hjust = 0.5)) + # set title to be at center
  xlab('Density') +
  ylab('Count')

rplot = ggplot(aes(x = density), data = wdf) + 
  geom_bar(stat = "count", fill =I('#2c7fb8'), width = 0.0001) + # bar polt of density
  scale_x_continuous(limit = c(0.99, 1.0), breaks = seq(0.99, 1.0, by = 0.0025)) + 
  ggtitle("Count of Wine by Density Levels (Revised Plot)") + 
  theme(plot.title = element_text(hjust = 0.5)) + # set title to be at center
  xlab('Density') +
  ylab('Count')

grid.arrange(bplot, rplot)

```

**The basic plot of the density shows that the wine records is ranged \
from 0.98 to 1.039 density levels with mean value of (0.9940). After several \
adjustments to the x-axis scale, most of the wines are ranged \
from 0.9910 to 0.9980 density level**

###***PH:***

```{r message=F, warning=F}

summary(wdf$pH) # summary of density

bplot = ggplot(aes(x = pH), data = wdf) + 
  geom_bar(stat = "count", fill =I('#2c7fb8')) + # bar polt of PH
  ggtitle("Count of Wine by PH Levels (Basic Plot)") + 
  theme(plot.title = element_text(hjust = 0.5)) + # set title to be at center
  xlab('PH') +
  ylab('Count')

rplot = ggplot(aes(x = pH), data = wdf) + 
  geom_bar(stat = "count", fill =I('#2c7fb8')) + # bar polt of PH
  scale_x_continuous(limit = c(2.8, 3.6), breaks = seq(2.8, 3.6, by = 0.05)) + 
  ggtitle("Count of Wine by PH Levels (Revised Plot)") + 
  theme(plot.title = element_text(hjust = 0.5)) + # set title to be at center
  xlab('PH') +
  ylab('Count')

grid.arrange(bplot, rplot)

```

**The basic plot of the PH shows that the wine records is ranged \
from 2.72 to 3.82 density levels with mean value of (3.188). After several\
adjustments to the x-axis scale, most of the wines are ranged \
from 2.9 to 3.45 PH level**

###***Sulphates:***

```{r message=F, warning=F}

summary(wdf$sulphates) # summary of Sulphates

bplot = ggplot(aes(x = sulphates), data = wdf) + 
  geom_bar(stat = "count", fill =I('#2c7fb8')) + # bar polt of sulphates
  ggtitle("Count of Wine by Sulphates Levels (Basic Plot)") + 
  theme(plot.title = element_text(hjust = 0.5)) + # set title to be at center
  xlab('Sulphates') +
  ylab('Count')

rplot = ggplot(aes(x = sulphates), data = wdf) + 
  geom_bar(stat = "count", fill =I('#2c7fb8')) + # bar polt of sulphates
  scale_x_continuous(limit = c(0.25, 0.85), breaks = seq(0.25, 0.75, by = 0.05)) + 
  ggtitle("Count of Wine by Sulphates Levels (Revised Plot)") + 
  theme(plot.title = element_text(hjust = 0.5)) + # set title to be at center
  xlab('Sulphates') +
  ylab('Count')

grid.arrange(bplot, rplot)

```

**The basic plot of the Sulphates shows that the wine records is ranged \
from 0.22 to 1.08 density levels with mean value of (0.4898). After several \
adjustments to the x-axis scale, most of the wines are ranged \
from 0.30 to 0.70 Sulphates level**

###***Alcohol:***

```{r message=F, warning=F}

summary(wdf$alcohol) # summary of alcohol

bplot = ggplot(aes(x = alcohol), data = wdf) + 
  geom_bar(stat = "count", fill =I('#2c7fb8'), width = 0.1) + # bar polt of alcohol
  ggtitle("Count of Wine by Alcohol Levels (Basic Plot)") + 
  theme(plot.title = element_text(hjust = 0.5)) + # set title to be at center
  xlab('Alcohol') +
  ylab('Count')

rplot = ggplot(aes(x = alcohol), data = wdf) + 
  geom_bar(stat = "count", fill =I('#2c7fb8'),width = 0.1) + # bar polt of alcohol
  scale_x_continuous(limit = c(8.5, 12.0), breaks = seq(8.0, 14, by = 0.25)) + 
  ggtitle("Count of Wine by Alcohol Levels (Revised Plot)") + 
  theme(plot.title = element_text(hjust = 0.5)) + # set title to be at center
  xlab('Alcohol') +
  ylab('Count')

grid.arrange(bplot, rplot)

```

**The basic plot of the Alcohol shows that the wine records is ranged \
from 8 to 14.20 alcohol levels with mean value of (10.51). After several \
adjustments to the x-axis scale, most of the wines are ranged \
from 8.6 to 11.8 alcohol level**


# Univariate Analysis

***

### What is the structure of your dataset?

**The dataset explored contains (4898) white wine quality records of 13 variables for each record**

***

### Description of dataset attributes: 

***

Attribute Number| Attribute| Description 
----|-------------|----------------------------------------------------------------- 
1|X|Record number in the dataset        
2| fixed acidity| Most acids involved with wine or fixed or nonvolatile \
(do not evaporate readily)
3| volatile acidity|  The amount of acetic acid in wine, which at too high of \
levels can lead to an unpleasant, vinegar taste
4| citric acid        | Found in small quantities, citric acid can add \
'freshness' and flavor to wines
5| residual sugar    | The amount of sugar remaining after fermentation stops,\
it's rare to find wines with less than 1 gram/liter and wines with greater than\
45 grams/liter are considered sweet
6| chlorides        | The amount of salt in the wine
7| free sulfur dioxide | The free form of SO2 exists in equilibrium between \
molecular SO2 (as a dissolved gas) and bisulfite ion; it prevents microbial \
growth and the oxidation of wine
8|  total sulfur dioxide| Amount of free and bound forms of S02; in low\
concentrations, SO2 is mostly undetectable in wine, but at free SO2 \
concentrations over 50 ppm, SO2 becomes evident in the nose and taste of wine
9| Density |The density of water is close to that of water depending on the \
percent alcohol and sugar content
10| pH | Describes how acidic or basic a wine is on a scale from 0 \
(very acidic) to 14 (very basic); most wines are between 3-4 on the pH scale
11| sulphates | A wine additive which can contribute to sulfur dioxide gas \
(S02) levels, wich acts as an antimicrobial and antioxidant
12| alcohol | The percent alcohol content of the wine
13| quality | (score between 0 and 10)

***

### What is/are the main feature(s) of interest in your dataset?

**The main feature in the dataset that i think will help supporting my \
investigation is Quality:**

**The quality in our dataset is the score given to record of wine (0 - 10)\
where 0 is very bad and 10 is very excellent. It is the main feature of \
investigation since the rating has been given by at least 3 wine experts**

***

### What other features in the dataset do you think will help support your \
investigation into your feature(s) of interest?

**Other features of investigations will the characterstics of wine compared\
with qualilty which are:**

**Feature**|**Comment**
-------------|-----------------------------------------------------------------
**Residual Sugar**|Residual sugar determines how sweet the wine is. It will \
be a good investigation area to see what residual sugar levels are rated as \
high quality
**Volatile Acidity**|I assume that the amount of acetic acid is a good quality\
indicator which is subject to investigation.
**PH**|Acidity levels can also be a good quality indicator, here in our dataset\
the pH levels scales from 0 (very acidic) to 14 (very basic)
**Alcohol**|Since wine is alcoholic beverage, the % of alcohol will be a good\
quality indicator as well
***
 

### Did you create any new variables from existing variables in the dataset?

**Yes I did create (2) new variables to store the rating levels into \
(quality.rating).The other new variable is (rating) to categorize ratings\
as High[8-9], Meduim[5-7], and Low[3-4].**

***

### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form \
of the data? If so, why did you do this?

**The residual sugar and total sulfur dioxide looked unsual which had a very \
wide range of distribution. Both plots were adjusted to show the majorirty of \
data distribution of each feature. The reason to perform such adjustement is to\
exclude outliers from the plots.**

***

# Bivariate Plots Section


```{r, message=F, warning=F, fig.width=10, fig.height=10,Bivariate_Plots}


#tmpdata <- wdf[ , !names(wdf) %in% c('X')]  # exlcuding the ID column (X) from the dataset

tmpdata <- subset(wdf, select = -c(X,quality.rating,rating)) # exlcuding the columns (X, quality.rating, and rating) from the dataset

#summary(tmpdata)

corrwdf <- subset(tmpdata, quality > 3 & quality <9) #excluding the ouliers (25 entries)

cor(corrwdf)

col <- colorRampPalette(c("#BB4444", "#EE9988", "#FFFFFF", "#77AADD", "#4477AA"))
correlate = cor(corrwdf)
corrplot(correlate, method="color", col=col(200),  
         diag=FALSE, 
         type="upper", order="hclust", 
         title="Correlation Matrix", 
         addCoef.col = "black", # Add coefficient of correlation
         insig = "blank", 
         number.cex = 1.5,
         tl.cex = 1.5,
         mar=c(0,0,1,0))

```

# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?

**My main feature of interest is the quality and from the correlation plot \
matrix, the only noticable positive correlcation between quality and other \
features is alcohol vs quality relationship. There is a positive correlation \
of 0.44 between quality and alcohol, which means that the more alcohol the \
better quality of wine.**


### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?

**From the correlation plot matrix, I can see that there is a very strong \
positive correlation of 0.84 betwen density and residual sugar. On the other\
hand, there is negative correlation of 0.78 between density and alcohol.**

### What was the strongest relationship you found?

**The strongest relationship between the features of white wine is found \
between density and residual sugar. The correlation between the two features \
is +0.84.**


# Multivariate Plots Section

###***Alcohol VS. Quality (Main Feature of Interest)***

```{r message=F, warning=F, Multivariate_Plots}

# Main feature of intereset plot

ggplot( aes( x= quality.rating, y = alcohol), data = subset(wdf, quality > 3 & quality <9)) + 
  geom_boxplot(alpha = 1)  +
  ggtitle("White Wine Quality by Alcohol Level") + 
  theme(plot.title = element_text(hjust = 0.5)) + # set title to be at center
  xlab('Quality Rating') +
  ylab('Alcohol')



```


###***Density VS. Residual Sugar (Other features of Interests)***

```{r message=F, warning=F}

# other features of interest plots

ggplot( aes( x= residual.sugar, y = density), data = subset(wdf, quality > 3 & quality <9)) + 
  geom_jitter(aes(color = rating), alpha = 1, size = 2) + 
  scale_color_brewer(type = 'div', palette = 4) +
  coord_cartesian(xlim = c(0, quantile(wdf$residual.sugar, 0.99)), 
                  ylim = c(quantile(wdf$density, 0.0001), quantile(wdf$density, 0.99))) +
  geom_smooth(method = lm,color = "black") +
  ggtitle("Density VS. Residual Sugar") + 
  theme(plot.title = element_text(hjust = 0.5),panel.background = element_rect(fill = 'wheat')) + # set title to be at center
  xlab('Residual Sugar') +
  ylab('Density')

```

###***Density VS. Alcohol (Other features of Interests)***

```{r message=F, warning=F}

# other features of interest plots

ggplot( aes( x= alcohol, y = density), data = subset(wdf, quality > 3 & quality <9)) + 
  geom_jitter(aes(color = rating), alpha = 1, size = 2) + 
  scale_color_brewer(type = 'div', palette = 4) +
  coord_cartesian(xlim = c(quantile(wdf$alcohol, 0.0001), 
                           quantile(wdf$alcohol, 0.99)), 
                  ylim = c(quantile(wdf$density, 0.0001), 
                           quantile(wdf$density, 0.99))) +
  geom_smooth(method = lm, color = "black") +
  ggtitle("Density VS. Alcohol") + 
  theme(plot.title = element_text(hjust = 0.5),panel.background = element_rect(fill = 'wheat')) + # set title to be at center
  xlab('Alcohol') +
  ylab('Density')

```

###***Linear Model Regression***

```{r  message=F, warning=F}

fit = lm(quality ~ alcohol + residual.sugar + density, data = wdf )
summary(fit)

```


# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?


**From the plot "White Wine Quality by Alcohol Level", I can see that the \
quality of wine gets better with alcohol level greater than 9.5 (for 5+ rating)**





### Were there any interesting or surprising interactions between features?


**Another area of interest was to see the relationship between density and \
residual sugar since there is a strong positive correlation between the \
two features.**

**From the plot "Density VS. Residual Sugar", there is a linear relationship\
between the density and residual sugar. The higher the density, there is higher \
sugar exist in white wine.**

**However, I have noticed a very intersting insight from the same plot which \
is that most of the high quality of wine has less density and hense\
less residual sugar.**

**As for the plot "Density VS. Alcohol", it is very clear that higher alcohol \
levels are associated with lower density levels.**

### OPTIONAL: Did you create any models with your dataset? Discuss the \
strengths and limitations of your model.

**Yes I did create a linear regression model to predict the quality of wine \
with multiple variables (alcohol, residual sugar, and density)**

**The R-squared value generated by the model is 0.21. This means that \
about 21% of wine quality is due to the levels of alcohol, residual sugar, \
and density. The generated model is weak and indicates that there are other\
factors that affect the quality of wine other than the 3 features \
tested in the model.**


------

# Final Plots and Summary


### Plot One
```{r echo=FALSE,fig.width=10, fig.height=5, Plot_One}

ggplot(aes(x = quality.rating), data = wdf) + 
  stat_count(width = 0.8, fill =I('#2c7fb8')) +
  facet_wrap(~rating, scales = "free" ) +
  ggtitle("Count of Wine Quality by Rating") + 
  theme(plot.title = element_text(hjust = 0.5)) + # set title to be at center
  xlab('Quality (0-10)') +
  ylab('Count')


```

### Description One

**The bar plot of quality shows that most of the quality is distributed in \
the Medium range (5-7 rating) while there are few observations that are \
rated as low or high.**

### Plot Two
```{r echo=FALSE, fig.width=10, fig.height=5, Plot_Two}

#corrplot(correlate, method="color", col=brewer.pal(n = 8, name = "PuOr"),  
#     diag=FALSE, 
#         type="upper", order="hclust", 
#         title="Correlation Matrix", 
#         addCoef.col = "black", # Add coefficient of correlation
#         insig = "blank", 
#         number.cex = 1.5,
#         tl.cex = 1.5,
#         mar=c(0,0,1,0))

ggplot( aes( x= quality.rating, y = alcohol,fill=rating), data = wdf) + 
  geom_boxplot(alpha=0.3)  +
  facet_wrap(~rating, scales = "free") +
  scale_fill_brewer(palette="Dark2") +
  ggtitle("White Wine Quality by Alcohol (% by volume)") + 
  theme(plot.title = element_text(hjust = 0.5)) + # set title to be at center
  xlab('Quality Rating (0-10)') +
  ylab('Alcohol (% by volume)')



```

### Description Two

**Looking at the main feature of interest (Quality) correlation with Alcohol,\
The plot shows that the ratings of wine is higher for higher alcohol level.\
The plot also explains the positive correlation of 0.44 between quality and\
alcohol**

### Plot Three
```{r echo=FALSE, fig.width=10, fig.height=8, Plot_Three}

ggplot( aes( x= residual.sugar, y = density), data = subset(wdf, quality > 3 & quality <9)) + 
  geom_jitter(aes(color = rating), alpha = 1, size = 2) + 
  scale_color_brewer(type = 'div', palette = 4) +
  coord_cartesian(xlim = c(0, quantile(wdf$residual.sugar, 0.99)), 
                  ylim = c(quantile(wdf$density, 0.0001), quantile(wdf$density, 0.99))) +
  geom_smooth(method = lm,color = "black") +
  ggtitle("Density VS. Residual Sugar") + 
  theme(plot.title = element_text(hjust = 0.5),panel.background = element_rect(fill = 'wheat')) + # set title to be at center
  xlab('Residual Sugar (g / dm^3)') +
  ylab('Density (g / cm^3)')

```

### Description Three

**The "Density VS. Residual Sugar" plot shows that there is a linear \
relationship between the density and residual sugar, which explains the \
strength of correlation factor between the two features (+0.84). The higher \
the density, the higher residual sugar exist in white wine.**

**In addition, a very intersting insight from the same plot that is most of\
the high quality of wine has less density and hense less residual sugar which \
means that sweetness in wine is not of a good quality feature.**

------

# Reflection

>**In this project, I have experienced the exploratory data analysis process \
by exploring the White Wine dataset which consist of 4898 observations\
and 13 variables.**
>
>**Exploration Findings:**
>
>>**The quality of wine varies from 3 to 9 in our dataset. There are no 0 or 10\
rated wine in the dataset, the minimum rated wine is 3 and the maximum is 9.\
The median quality is 6 and there are very few wine records that are rated as\
3 or 9 (25 records out of the whole dataset).**
>
>>**Most of the quality is distributed in the Medium range (5-7 rating) while \
there are few observations that are rated as low or high.**
>
>>**The most strong (postivie) correlation is between Density and Residual \
Sugar with a correlation of 0.84 while the most strong (negative) correlation is\
between Density and Alcohol with a correlation of 0.78.**
>
>>**Looking at the main feature of interest (Quality) correlation with other \
features, the most strong correlation relationship is with Alcohol which has a \
correlation of 0.44.**
>
>>**The R-squared value generated by the model is 0.21. This means that about \
21% of wine quality is due to the levels of alcohol, residual sugar, and \
density. The generated model is weak and indicates that there are other factors\
that affect the quality of wine other than the 3 features tested in the model.**
>
>**It was surprizing that the high quality of wine has less residual sugar which\
means that sweetness in wine is not of a good quality feature although I was \
thinking that the sweeter the better.**
>
>**I have learned a lot in this experience of exploration about visualising and\
summarizing data using R. I have also learned that right plot scaling makes huge\
differences when interpreting plots and summarizing findings although I did\
strugle at the begining when it comes to finding the right scales and getting\
the plots to show what I intended to.**
>
>**I think a future work to this dataset can be adding another feature that has\
a strong correlation with quality other than Alcohol which is the only \
correlatable feature (with quality) in this dataset. This would provide a\
deeper experience with correlation and coming up with more findings.**

>**Below References were very useful throughout the project work**
>
>[1- Visualize Correlation](http://www.sthda.com/english/wiki/visualize-corre\
lation-matrix-using-correlogram)
>
>[2- Corrplot Tilte Position](https://stackoverflow.com/questions/40509217/how-\
to-have-r-corrplot-title-position-correct)
>
>[3- Linear Models in R](http://data.princeton.edu/R/linearModels.html)
>
>[4- Regression](https://www.statmethods.net/stats/regression.html)
>
>[5- Linear Modeling in R](http://blog.yhat.com/posts/r-lm-summary.html)
